---
title: "SurvBal_Vignette"
output: rmdformats::readthedown
vignette: >
  %\VignetteIndexEntry{SurvBal_Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>"
)
```



## Overview

`SurvBal` enables the selection of microbiome balances in relation to censored survival or time-to-event outcomes, which are of considerable interest in many biomedical studies. The most commonly used survival models – the Cox proportional hazards and parametric survival (including accelerated failure time) models are included in the package, which are used in combination with step-wise selection procedures to identify the optimal associated balance of microbiome, i.e., the ratio of the geometric means of two groups of taxa’s relative abundances.

## Installation 

You can install `SurvBal` from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("yinglia/SurvBal")
```

## Getting Started

```{r, results=FALSE}
#loading the library
library(SurvBal)
```

You can load the sample data in `SurvBal` using the following code.

```{r}
#load example data
data("bacteria")
data("gvhd")
```

`bacteria` is a raw taxon count table with 63 samples and 139 taxa. `gvhd` is a survival object containing time and censoring/event information for each sample.


## Examples

### Select balance based on the Cox proportional hazards model

We can select the balance of microbiome based on the Cox proportional hazards Model by setting `model = "coxph"`. 


Here, we specify `mult_repl = TRUE` to use the geometric Bayesian multiplicative replacement method to impute the inflated zeros. If you would like to process the data by adding a small pseudo-count (0.5) to all raw counts instead, set `mult_repl = FALSE`.

We specify `selection_criterion = "min_decrement_pvalue"` so that the decrement of p-value along the forward selection path will be calculated, the balance before the first decrement that is smaller than selection_threshold (the default value is 0.15) will be selected. If “min_pvalue” is used, the balance that has the smallest p-value along the forward selection path will be selected. The default is “min_decrement_pvalue”.


```{r, results='hide'}
select_coxph = balance_selection(
  Surv_obj = gvhd,
  data = bacteria,
  mult_repl = TRUE,
  model = "coxph",
  selection_criterion = "min_decrement_pvalue"
)
```

We can find the selection path in `selection_path`:

```{r}
select_coxph$selection_path[,1:2]
```

We can print the final fitted Cox PH model for the survival outcome, `gvhd` , with the selected balance:

```{r}
select_coxph$survival_model
summary(select_coxph$survival_model)
```

We can check which taxa are included in the numerator of the selected balance, as well as taxa in the denominator of the selected balance:

```{r}
select_coxph$balance_name$numerator
select_coxph$balance_name$denominator
```

And here is the value of the final selected balance of microbiome for each sample:

```{r}
select_coxph$balance
```

Finally, we can visualize the survival plot with respect to several quantiles of the final selected balance of microbiome with `1-alpha` confidence interval. The quantiles of the final selected balance for plot are specified through the argument `quantile_plotted`, of which the default value is
`c(0.25, 0.75)`. The default value of alpha is `0.05`.

```{r}
select_coxph$survival_plot
```

From the plot, we can see that the 25th percentile (lower quartile) of the sample microbiome balance is -0.6735, and the 75th percentile (upper quartile) is 0.5512. The survival curves for the two corresponding values of the microbiome balance are shown in the plot.

### Select balance based on the parametric survival model

We can also select the balance based on the parametric survival model by setting `model = "parametric"`
, and specify the distribution via the argument `dist`.

```{r, results='hide'}
select_weibull = balance_selection(
  Surv_obj = gvhd,
  data = bacteria,
  mult_repl = FALSE,
  model = "parametric",
  dist = "weibull",
  selection_criterion = "min_decrement_pvalue"
)
```

Similarly, we can find the selection path in `selection_path`:

```{r}
select_weibull$selection_path[,1:2]
```

We can print the final fitted parametric survival model for the survival outcome, `gvhd`, with the selected balance:

```{r}
select_weibull$survival_model
```

We can check which taxa are included in the numerator of the selected balance, as well as taxa in the denominator of the selected balance:



```{r}
select_weibull$balance_name
```

And here is the value of the final selected balance of microbiome for each sample:

```{r}
select_weibull$balance
```

Finally, we can visualize the survival plot:

```{r}
select_weibull$survival_plot
```

